### 2018-01-19 includes/abstract.tex
Rewriting some sections and srestructuring paper. Removing old files.

    In this paper we present \fuzzer{}: an \smtlib{} problem fuzzer and generator for string solvers. We provide a rich repository of problems we generated with \fuzzer{} in \smt{} format. We also analyse \numSolvers{} mature string solvers: \theSolvers{}. We use \fuzzer{} to craft inputs that elicit poor performance in \us{} and \cvc{}, and analyse the causes of the performance degradations. We then describe an algorithm to mitigate the performance degradation in \us{}. Finally, we describe several bugs in \us{} that were exposed by \fuzzer{}.

>>>>

    In this paper we present \fuzzer{}: a modular \smtlib{} problem transformer and generator for string solvers. We supply a rich repository of problems we generated with \fuzzer{} in \smt{} format, and demonstrate its usefulness in the development of \us{} by presenting defects it helped find and fix. We also use \fuzzer{} to exercise \theSolvers{} with suites of problems that elicit poor performance in one solver but not the other. We analyse our experimental results to identify the causes of the performance degradations, and present our findings.

### 2018-01-19 includes/experiment.tex
Rewriting some sections and srestructuring paper. Removing old files.

    Here we present two particular sets of problems generated by \fuzzer{}: \cHard{} and \zHard{}. These two problem sets were unique in their ability to elicit poor performance in one solver, but not another. Full details of test scripts, configurations, and problems that were used in the experiment can be found in the Appendix.

    % TODO:
    % - supply machine environment description
    % - supply install scripts that produce trace builds
    % - supply test scripts that invoke the solvers and produce traces

    \subsection{Data}

        \subsubsection{Performance on \cHard{}}

            On this problem set \cvc{} did poorly, but \us{} did well. The graphs in Figures~\ref{fig:z3str3-cvc4-hard} and ~\ref{fig:cvc4-cvc4-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-cvc4-hard.png}
                    \caption{Performance of \cvc{} on \cHard{}}
                    \label{fig:cvc4-cvc4-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-cvc4-hard.png}
                    \caption{Performance of \us{} on \cHard{}}
                    \label{fig:z3str3-cvc4-hard}
                \end{subfigure}
                \caption{Results on \cHard}
                \label{fig:cvc4-hard}
            \end{figure}

            The \cvc{} trace file indicates that it performs the following operations:

@@

            At this stage it discovers that there is a conflict in one of its equivalence classes, as shown in Figure~\ref{fig:cvc4-conflict}. After discovering the conflict, the rest of the trace is filled with apparently the same operation repeated hundreds of thousands of times. After this operation is finished, \cvc{} recalls the earlier conflict and returns \texttt{unsat}.

@@

            \us{} showed different behaviour on this problem set. It follows a similar process as \cvc{}, except it does not flatten concatenations. In contrast to \cvc{}, \us{} immediately returns \texttt{unsat} as soon as it verifies consistency of equivalence classes as it constructs them, as shown in Figure~\ref{fig:z3str3-conflict}.

@@

            % TODO:
            % - \us{} cactus plot
            % - \cvc{} cactus plot
            % - comparison plot

        \subsubsection{Performance on \zHard{}}

            On this problem set \cvc{} did well, but \us{} did poorly. The graphs in Figures~\ref{fig:z3str3-z3str3-hard} and ~\ref{fig:cvc4-z3str3-hard} show the performance results for both solvers. We produce and analyse the execution trace files both solvers' execution on the problem set, and present them in the Appendix.

            \begin{figure}[h]
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/cvc4-z3str3-hard.png}
                    \caption{Performance of \cvc{} on \zHard{}}
                    \label{fig:cvc4-z3str3-hard}
                \end{subfigure}
                \begin{subfigure}{.5\textwidth}
                    \includegraphics[width=\textwidth]{data/graphs/z3str3-z3str3-hard.png}
                    \caption{Performance of \us{} on \zHard{}}
                    \label{fig:z3str3-z3str3-hard}
                \end{subfigure}
                \caption{Results on \zHard}
                \label{fig:z3str3-hard}
            \end{figure}

            Both solvers exhibit similar patterns in their trace files. They differ in their strategies for assigning possible satisfying assignments. \cvc{} assigns the first variable to a constant and all other variables to the empty string. \us{} however seems to ``drag'' the constant across the equation, splitting its characters among the variables in it.

            % TODO:
            % - \us{} cactus plot
            % - \cvc{} cactus plot
            % - comparison plot

        \subsubsection{Performance Concats-Extracts}

            \todo{}: No data on this yet.

    \subsection{Analysis}

        This section presents our analysis of the observed behaviour of \cvc{} and \us{} on \cHard{} and \zHard{} problems.

        \todo{}: On \cHard{}, \cvc{} does poorly and \us{} does well because ...

        \todo{}: On \zHard{}, \cvc{} does well and \us{} does poorly because ...

>>>>

    We generated several problem suites with \fuzzer{} that made one solver perform poorly, but not another. We call those suites: \cHard{} and \zHard{}.

    On \cHard{} \cvc{} did poorly, but \us{} did well. The graphs in Figure~\ref{fig:cvc4-hard} show the performance results for both solvers.

    On \zHard{} \cvc{} did well, but \us{} did poorly. The graphs in Figure~\ref{fig:z3str3-hard} show the performance results for both solvers.

    \todo{Add data on Concats-Extracts}

    On Concats-Extracts ...

    \begin{figure}[H]
        \begin{subfigure}{.5\textwidth}
            \includegraphics[width=\textwidth]{data/graphs/cvc4-cvc4-hard.png}
            \caption{Performance of \cvc{} on \cHard{}}
            \label{fig:cvc4-cvc4-hard}
        \end{subfigure}
        \begin{subfigure}{.5\textwidth}
            \includegraphics[width=\textwidth]{data/graphs/z3str3-cvc4-hard.png}
            \caption{Performance of \us{} on \cHard{}}
            \label{fig:z3str3-cvc4-hard}
        \end{subfigure}
        \caption{Results on \cHard}
        \label{fig:cvc4-hard}
    \end{figure}

    \begin{figure}[H]
        \begin{subfigure}{.5\textwidth}
            \includegraphics[width=\textwidth]{data/graphs/cvc4-z3str3-hard.png}
            \caption{Performance of \cvc{} on \zHard{}}
            \label{fig:cvc4-z3str3-hard}
        \end{subfigure}
        \begin{subfigure}{.5\textwidth}
            \includegraphics[width=\textwidth]{data/graphs/z3str3-z3str3-hard.png}
            \caption{Performance of \us{} on \zHard{}}
            \label{fig:z3str3-z3str3-hard}
        \end{subfigure}
        \caption{Results on \zHard}
        \label{fig:z3str3-hard}
    \end{figure}

\section{Analyses}

    In this section, we analyse the experimental results from the previous section. Although it is satisfying to find a class of problems that makes only one solver behave a certain way, it is only practically useful when the behaviour's cause is also understood.

    \todo{Maybe organise these by solver, instead of by problem?}

    \subsection{\cHard{}}

        The \cvc{} trace file suggests that it performs the following operations:

@@

        At this stage it discovers that there is a conflict in one of its equivalence classes, as shown in Figure~\ref{fig:cvc4-conflict}. After discovering the conflict, the rest of the trace is filled with apparently the same operation repeated hundreds of thousands of times. This operation dominates the content of the trace. After this expensive operation is finished, \cvc{} appears to recall the earlier conflict and returns \texttt{unsat}.

@@

        \us{} behaves differently. It follows a similar process as \cvc{}, except it does not flatten concatenations. In contrast to \cvc{}, \us{} immediately returns \texttt{unsat} as soon as it verifies consistency of equivalence classes as it constructs them, as shown in Figure~\ref{fig:z3str3-conflict}.

@@

        We were not equipped well enough to debug this behaviour in \cvc{}, but we intend this result to be useful to its authors in improving it.

    \subsection{\zHard{}}

        Both solvers exhibit similar patterns in their trace files. They differ in their strategies for assigning possible satisfying assignments.

        \todo{Explain why \us{} performs poorly.}

### 2018-01-19 includes/introduction.tex
Rewriting some sections and srestructuring paper. Removing old files.

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and practically useful. Moreover a standard input language, \smt{}, has been developed and many mature solvers support it.

    In this paper, we present the following contributions:

    \begin{enumerate}
        \item \fuzzer{}: a fuzzer of \smt{} string and regex problems.
        \item Generated inputs that elicit poor performance in \us{} and \cvc{}.
        \item A recommended fix to \us{} that alleviates the poor performance.
        \item A large suite of \smt{} string and regex problems we generated with \fuzzer{}.
    \end{enumerate}

>>>>

    Many SMT solvers that support the theory of strings have been developed over the past decade, and several have matured to the point of being performant and useful in practice. \theSolvers{}\cite{cvc4}\cite{z3str3} are \numSolvers{} such solvers. Moreover a standard input language for problem instances, \smt{}\cite{smt}, has been developed and many solvers support it. To validate their solvers, many authors have relied on hand-crafted input suites or real-world examples from industrial applications \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. Indeed several suites of industrial and hand-crafted problems have been compiled \cite{kaluza} \cite{kausler} to both verify solvers and benchmark their performance. However though these suites help solver developers debug and verify their solvers, they do not age well. As solvers grow more complex and expressive, fixed benchmarks fail to exercise new behaviours. As a result, more and more solver functionality goes untested or unverified.

    We introduce a modular fuzzer for string solvers, \fuzzer{}, and describe its utility as an exploratory verification tool. We show its effectiveness by presenting defects we were able to find in existing solvers.

    \subsection{Contributions}

    \begin{enumerate}
        \item \textbf{The \fuzzer{} tool}: We describe a modular fuzzer which can transform and generate \smt{} string and regex problems. We briefly document its components, modular architecture, and usage patterns to demonstrate its usefulness as an exploratory tool for validating not only solver implementations, but also algorithmic assumptions.
        \item \textbf{A suite of \smt{} problems}: We present a suite of \smt{} string and regex problems we generated with \fuzzer{}. We justify the existence of each suite by a motivating example from the development of \us{}. We reveal specific defects that this suite helped uncover and fix, as well as specific conjectures that it helped validate or disprove.
        \item \textbf{Experimental Results}: We present problem classes we found that made one solver perform poorly, but not another. We show differences in solver behaviour on the suites \cHard{} and \zHard{}.
        \item \textbf{Analysis}: We analyse our experimental results by inspecting the execution traces of the solvers on the given problems. Where possible, we explain our observations. For \cvc{} we draw superficial conclusions based solely on the execution traces, and invite its authors to join us in analysing them. For \us{} we pinpoint algorithmic shortcomings that cause the poor performance.
    \end{enumerate}

### 2018-01-19 includes/prior.tex
Rewriting some sections and srestructuring paper. Removing old files.

\section{Prior Work}

    There are several fuzzers and problem generators currently available, but none of them support string or regex generation or transformation. For example the FuzzSMT\cite{fuzzsmt} tool generates \smt{} problems with bit vectors and arrays, but does not support strings or regexes. The SMTpp\cite{smtpp} tool preprocesses and simplifies problems, but does not generate new ones or fuzz existing ones.

@@

    \hfill \\
    \todo{}: references

>>>>

\section{Related Work}

    Naturally, many solver developers author their own unit test suites to verify their solvers \cite{cvc4-tests} \cite{z3str3-tests} \cite{z3str2-tests}. In addition, several popular problem suites are publicly available for solver verification, such as the Kaluza \cite{kaluza} and Kausler \cite{kausler} suites.

    There are likewise several fuzzers and problem generators currently available, but none of them support string or regex generation or transformation. For example the FuzzSMT\cite{fuzzsmt} tool generates \smt{} problems with bit vectors and arrays, but does not support strings or regexes. The SMTpp\cite{smtpp} tool preprocesses and simplifies problems, but does not generate new ones or transform existing ones.

@@

### 2018-01-19 includes/problems.tex
Rewriting some sections and srestructuring paper. Removing old files.

    \subsection{\fuzzer{}-Generated}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Lengths
            \item Lengths-Concats
            \item Concats
            \item Concats-Extracts
            \item Overlaps
            \item Regex
            \item Many-Regexes
            \item Regex-Deep
            \item Regex-Pair
            \item Regex-Lengths
            \item Random??
        \end{itemize}

        Most of these problem sets either have a dedicated generator in \generator{}, or are variations of one. While it is likely that some of the pre-made suites will be immediately useful in analysing a solver, we also found that mass randomisation was helpful in finding some bugs, and we encourage this pattern as well. \generator{} is designed to be used in an exploratory fashion. For example the following usage has helped us find bugs in \us{}:

\begin{verbatim}
while true; do stringfuzzg --random regex | z3str3 -in; done
\end{verbatim}

    \subsection{\fuzzer{}-Transformed}

        This suite contains the following problem sets:

        \begin{itemize}
            \item Amazon??
            \item Regex??
        \end{itemize}

        Most of these problem sets were generated by using one root problem and transforming it with \transformer{} into many similar problems.

    \subsection{Unit Tests}

        This suite covers corner cases of solver behaviour. We have used the instances in this suite to find bugs in \us{}.

    \subsection{Bugs Found}

@@

        \begin{enumerate}
            \item length lookup using eqc root
            \item length-testing bug
        \end{enumerate}

>>>>

    \subsubsection{\fuzzer{}-Generated}

        Most of the problem sets in this suite either have a dedicated generator in \generator{}, or are variations of one. This suite contains the following problem sets: \\

        \todo{Fill in descriptions.}

        \begin{tabular}{|c|c|}
            \hline
            \textbf{Name}    & \textbf{Description} \\ \hline
            Lengths          & \\ \hline
            Lengths-Concats  & \\ \hline
            Concats          & \\ \hline
            Concats-Extracts & \\ \hline
            Overlaps         & \\ \hline
            Regex            & \\ \hline
            Many-Regexes     & \\ \hline
            Regex-Deep       & \\ \hline
            Regex-Pair       & \\ \hline
            Regex-Lengths    & \\ \hline
        \end{tabular}

    \subsubsection{\fuzzer{}-Transformed}

        Most of the problem sets in this suite were generated by using one root problem and transforming it with \transformer{} into many similar problems. This suite contains the following problem sets: \\

        \todo{Fill in descriptions.}

        \begin{tabular}{|c|c|}
            \hline
            \textbf{Name} & \textbf{Description} \\ \hline
            Amazon        & \\ \hline
            Regex         & \\ \hline
        \end{tabular}

\subsection{Results}

@@

    \todo{Describe the length lookup bug and length-testing bug, giving the before/after commits.}

    While it is likely that some of the pre-generated suites will be immediately useful in analysing a solver, we also found that randomisation was helpful in exploratory debugging, and we encourage this pattern as well. For example the following usage can feed random problems to \cvc{} until it returns an error:

    {\scriptsize\begin{verbatim}
while stringfuzzg random | cvc4 --lang smt2.5 --tlimit=5000; do sleep 0; done\end{verbatim}}

    Some simple UNIX scripting can yield other interesting results. For example the following script finds a regex problem on which \us{} times out after 5 seconds:

    {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat hard-problem.smt2
        break
    fi
done\end{verbatim}}

### 2018-01-19 includes/stringfuzz.tex
Rewriting some sections and srestructuring paper. Removing old files.

    This section describes the fuzzer \fuzzer{}, which generates new \smt{} problems or transforms existing ones. \fuzzer{} is implemented as a Python package, and comes with several executables to generate, transform, and analyse \smt{} string problems. Its source code repository is available at \sourceRepo{}. It can either be installed from source, or from the Python PIP package repository.

    The \fuzzer{} code is organised to be easily extended: all generators and transformers are independent modules. It is also implemented as a Unix "filter", to enable easy integration with other tools (including itself). For example, the outputs of generators can be piped directly into transformers, and transformers can be chained to produce a tuned stream of challenging inputs to a solver.

    Below are descriptions of the tools that come with \fuzzer{}:

@@

@@

        \item[\texttt{stringfuzzx}] \hfill \\
            This tool transforms \smt{} problems.. It supports several transformers and options that control its output and input. The transformers it implements are:\\

@@

>>>>

    This section describes the fuzzer \fuzzer{}, which generates new \smt{} problems and transforms existing ones. \fuzzer{} is implemented as a Python package, and comes with several executables to generate, transform, and analyse \smt{} string problems.

    % Its source code repository is available at \sourceRepo{}. It can either be installed from source, or from the Python PIP package repository.

    The \fuzzer{} code is organised to be easily extended: all generators and transformers are independent modules. It is also implemented as a Unix ``filter'', to enable easy integration with other tools (including itself). For example, the outputs of generators can be piped directly into transformers, and transformers can be chained to produce a tuned stream of challenging inputs to a solver.

    Below are descriptions of the tools that constitute \fuzzer{}:

@@

            \todo{Fill in descriptions.}

@@

        \item[\texttt{stringfuzzx}] \hfill \\
            This tool transforms \smt{} problems.. It supports several transformers and options that control its output and input. The transformers it implements are: \\

            \todo{Fill in descriptions.}

@@

            \todo{Add details about merging strategies.}
